\chapter*{ABSTRACT}
\begin{flushleft}
\textbf{Title:} Prototype development of the control unit for the X band transmission system\\
\textbf{Author:} Rodrigo Laurinovičs\\
\textbf{Supervising tutor:} Mg. Sc. Ing. Mārcis Bleiders\\
\textbf{Scope of work:} 38 pages, 5 tables, 51 figures, 32 bibliographical references, 0 appendices.
\textbf{Keywords:} X-band transmitter bias-up and bias-down, True RMS power detectors, TCP/IP, satellite communications
\\In the bachelor's thesis, a functional prototype for monitoring and biasing a high-power amplifier has been developed. The prototype is network-controllable and applicable for transmitter control in satellite communications in the X-band.\\

The paper describes the RT16 X-band system, the spectrum allocation of the Deep Space Network, existing control systems for power amplifiers, and power detectors. The achievable parameters and functionality required for the efficient use of transmitters in the existing Irbene radio telescope system have been defined.\\

To facilitate prototyping, each subsystem was developed as an independent module using modern, easily accessible electronic components and materials. Computer-aided design tools were used in the development of the module schematics.\\

Measurements of the key electrical parameters were performed on the manufactured subsystems. The prototype was tested in a transmission system, with performance-related parameter measurements conducted, achieved functionality verified, and conclusions drawn.
\end{flushleft}
